---
title: "TEST TEST 2 vireo"
format: 
  dashboard:
    orientation: rows
    embed-resources: true
params: 
    param_plot_list_grob: "C:/Users/Danne/R_projects/hub_cubi_cusco/output/hub_10/plot_list_grob.rds"
    param_ggextra_plot_list_grob: "C:/Users/Danne/R_projects/hub_cubi_cusco/output/hub_10/plotified_plot_list_grob.rds"
---


```{r setup}
#| warning: false
#| echo: false

library(tidyverse)
library(here)
library(grid)
library(gt)
library(purrr)

#plot_list_grob <-  read_rds(here("output", "hub_10","plot_list_grob.rds"))
plot_list_grob <- read_rds(params$param_plot_list_grob)
#ggextra_plot_list_cellbender_cells <- read_rds(here("output", "hub_10","plot_list_grob.rds"))
#clean_plot_list_cellbender_cells <-  read_rds(here("output", "hub_10","clean_plot_list.rds"))
```


```{r }
#plotified_ggextra_plot_list_cellbender_cells <-   read_rds(here("output", "hub_10","plotified_plot_list_grob.rds"))
plotified_ggextra_plot_list_cellbender_cells <- read_rds(params$param_ggextra_plot_list_grob)

```

#  {.sidebar}

**Data filtered by:**
```{r}
#grid::grid.newpage()
#grid.draw(plot_list_grob$text_filter_annotations, )
# Named list of character vectors



plot_list_grob$text_filter_annotations

```


# Page 1

## Row {height="50%"}

### Column {.tabset}

```{r}
#| title: Cellranger RNA, RNA clusters


grid::grid.newpage()
grid.draw(plot_list_grob$RNA_dimplot_cellranger)

```

```{r}
#| title: Cellranger RNA, vireo clusters


grid::grid.newpage()
grid.draw(plot_list_grob$RNA_dimplot_vireo)

```

```{r}
#| title: Cellranger RNA, cellbender clusters

grid::grid.newpage()
grid.draw(plot_list_grob$RNA_dimplot_cellbender)

```

### Column {.tabset}

```{r}
#| title: Cellbender RNA, cellranger clusters


grid::grid.newpage()
grid.draw(plot_list_grob$cellbender_RNA_dimplot_cellranger)

```

```{r}
#| title: Cellbender RNA, cellbender clusters


grid::grid.newpage()
grid.draw(plot_list_grob$cellbender_RNA_dimplot_cellbender)

```

```{r}
#| title: Cellbender RNA, vireo clusters


grid::grid.newpage()
grid.draw(plot_list_grob$cellbender_RNA_dimplot_vireo)

```





```{r}
#| title: Vireo SNV demux, cellranger clusters

grid::grid.newpage()
grid.draw(plot_list_grob$hashtag_dimplot_cellranger)

```



```{r}
#| title: Vireo SNV demux, cellbender clusters


grid::grid.newpage()
grid.draw(plot_list_grob$hashtag_dimplot_cellbender)

```

```{r}
#| title: Vireo SNV demux, vireo clusters


grid::grid.newpage()
grid.draw(plot_list_grob$hashtag_dimplot_vireo)

```




## Row

### Column {.tabset}

```{r}
#| title: knee nCount_RNA - vireo donor


grid::grid.newpage()
grid.draw(plotified_ggextra_plot_list_cellbender_cells$nCount_RNA_rank__vs__RNA_rank)

```

```{r}
#| title: knee nCount_RNA - vireo cell

grid::grid.newpage()
grid.draw(plotified_ggextra_plot_list_cellbender_cells[["nCount_RNA_rank__vs__RNA_rank_no_NA"]] )
```

```{r}
#| title: knee, marginal nCount_RNA

grid::grid.newpage()
grid.draw(plotified_ggextra_plot_list_cellbender_cells$nCount_RNA_rank__vs__RNA_rank_color_cellbender)
```

```{r}
#| title: cellbender, vireo, cellranger counts

grid::grid.newpage()
grid.draw(plot_list_grob$vireo_calls_barchart_grid_cellbender_vireo)

```

```{r}
#| title: cellbender prob.

grid::grid.newpage()
grid.draw(plot_list_grob$cellbender_prob__vs__RNA_rank_color_cellbender)

```

```{r}
#| title: cellbender histo

grid::grid.newpage()
grid.draw(plot_list_grob$cellbender_prob_histogram)

```


### Column

lets write some text herethat describes the filtering process and qc observations.

lets write some text herethat describes the filtering process and qc observations.
lets write some text herethat describes the filtering process and qc observations.
lets write some text herethat describes the filtering process and qc observations.
lets write some text herethat describes the filtering process and qc observations.
lets write some text herethat describes the filtering process and qc observations.
lets write some text herethat describes the filtering process and qc observations.


